A Phenomenological Model for the Intracluster Medium that matches X-ray and Sunyaev-Zel'dovich observations

### 2013-09-29 paper1.tex
conclusions revised

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
  {\bf Cosmological hydrodynamical simulations of galaxy clusters are still
    challenged to produce a model for the intracluster medium that matches all
    aspects of current X-ray and Sunyaev-Zel'dovich observations. To facilitate
    such comparisons with future simulations and to enable realistic cluster
    population studies for modeling e.g., non-thermal emission processes, we
    construct a phenomenological model for the intracluster medium that is based
    on a representative sample of X-ray clusters. We create a mock galaxy
    cluster catalog by assigning our gas density model to each cluster halo of
    the large collisionless $N$-body simulation MultiDark. Our clusters
    are identified with cool-core and non cool-core clusters according to a
    dynamical disturbance parameter. We demonstrate that our gas model matches
    the various observed Sunyaev-Zel'dovich and X-ray scaling relations as well
    as the X-ray luminosity function, thus enabling a reliable mock catalog for
    present and future instruments.  In a companion paper, we apply our catalogs
    to calculate non-thermal radio and gamma-ray emission of galaxy clusters. We
    make our cosmologically complete multi-frequency mock catalogs for the
    (non-)thermal cluster emission at different redshifts publicly and freely
    available online through the MultiDark database.}
\end{abstract}

@@

We note that our mock catalogs are not appropriate for all purposes because they
represent the average cluster population. While they are very useful as baseline
for new simulations, for comparisons to new observational samples, and for
non-thermal modeling of clusters, they are not appropriate to perform detailed
cosmology studies. Careful modeling of the response function of the considered
instrument in combination with light-cone mock catalogs would be required in this
case. Nevertheless, for volume-complete surveys, our mock catalogs will
certainly serve as valuable tools.  }

@@

{\bf We build a complete cosmological sample of galaxy clusters from the
  MultiDark $N$-body simulation with redshifts ranging from $z = 0$ to 1 and
  construct a \emph{phenomenological} model for the ICM. This is characterized
  by X-ray-inferred (CCC and NCCC) gas profiles (taken from the REXCESS sample)
  and a cluster mass-dependent gas fraction. We assign such a (cluster
  mass-dependent) gas density profile to each DM halo in our sample and sort it
  into the NCCC/CCC populations according to a dynamical disturbance parameter
  that is calculated from the DM distribution.  We obtain cosmologically
  complete mock catalogs of galaxy clusters that matches the observed
  $L_{\rmn{X, bol}}-M_{500}$, $Y_{\rmn{X}}-M_{500}$, and $Y_{\rmn{SZ}}-M_{500}$
  relations, as well as the X-ray luminosity function.  In the companion
  Paper~II, we additionally present a model for the radio and gamma-ray emission
  as a result of CR hadronic interactions with the ICM.

  We provide a cosmologically complete multi-frequency mock catalog for the
  (non-)thermal cluster emission at different redshifts. We make these catalogs
  publicly and freely available on-line through the MultiDark database
  (www.multidark.org). Our mock catalogs are good representations of the average
  cluster population. They allow quantitative comparisons of future
  hydrodynamical simulations, the modeling of non-thermal cluster emission from
  radio to gamma-rays, and the construction of cluster mock catalogs of the
  (non-)thermal cluster emission. We emphasize that the catalogs are not suited
  for detailed cosmological studies in their (present) published form, which do
  not address any survey response functions nor provide light-cone mock
  catalogs. However, future surveys should be able to test the underlying assumptions
  of our modeling approach, which is based on X-ray observations of clusters at
  low redshift ($z<0.2$) and adopts self-similar redshift evolution for
  extrapolating to higher redshifts. We hope that the community can make
  valuable use of these catalogs in synergy with the future radio, X-ray and
  gamma-ray data.}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
  {\bf Cosmological hydrodynamical simulations of galaxy clusters are still
    challenged to produce a model for the intracluster medium that matches all
    aspects of current X-ray and Sunyaev-Zel'dovich observations. To facilitate
    such comparisons with future simulations and to enable realistic cluster
    population studies for modeling e.g., non-thermal emission processes, we
    construct a phenomenological model for the intracluster medium that is based
    on a representative sample of X-ray clusters. We create a mock galaxy
    cluster catalog by assigning our gas density model to each cluster halo of
    the large collisionless $N$-body simulation MultiDark. Our clusters are
    identified with cool-core and non cool-core clusters according to a
    dynamical disturbance parameter. We demonstrate that our gas model matches
    the various observed Sunyaev-Zel'dovich and X-ray scaling relations as well
    as the X-ray luminosity function, thus enabling a reliable mock catalog for
    present surveys and forecasts for future instruments. In a companion paper,
    we apply our catalogs to calculate non-thermal radio and gamma-ray emission
    of galaxy clusters. We make our cosmologically complete multi-frequency mock
    catalogs for the (non-)thermal cluster emission at different redshifts
    publicly and freely available online through the MultiDark database.}
\end{abstract}

@@

We note that our mock catalogs are not appropriate for all purposes because the
density profile shapes that we adopt in our model represent the average cluster
population of cool-core and non cool-core systems (while we account for scatter
in the gas mass fractions, i.e., the normalization of the profiles). As such,
the models are very useful as a baseline for comparisons to new hydrodynamical
simulations and to new observational samples as well as for non-thermal modeling
of clusters. On the contrary, they are not appropriate to perform detailed
cosmology studies. Careful modeling of the response function of the considered
instrument in combination with light-cone mock catalogs would be required in
this case. Nevertheless, for volume-complete surveys, our mock catalogs will
certainly serve as valuable tools.  }

@@

{\bf We build a complete cosmological sample of galaxy clusters from the
  MultiDark $N$-body simulation with redshifts ranging from $z = 0$ to 1 and
  construct a \emph{phenomenological} model for the ICM. This is characterized
  by X-ray-inferred (CCC and NCCC) gas profiles (taken from the REXCESS sample)
  and a cluster mass-dependent gas fraction. We assign such a (cluster
  mass-dependent) gas density profile to each DM halo in our sample and sort it
  into the NCCC/CCC populations according to a dynamical disturbance parameter
  that is calculated from the DM distribution. Applying the model to our sample
  of cluster halos, we obtain volume complete mock catalogs for the
  thermal (SZ and X-ray) cluster emission. Our mock catalogs match the observed
  $L_{\rmn{X, bol}}-M_{500}$, $Y_{\rmn{X}}-M_{500}$, and $Y_{\rmn{SZ}}-M_{500}$
  relations, as well as the X-ray luminosity function.  In the companion
  Paper~II, we additionally present a model for the radio and gamma-ray emission
  as a result of CR hadronic interactions with the ICM.

  We provide cosmologically complete multi-frequency mock catalogs for the
  (non-)thermal cluster emission at different redshifts and make these catalogs
  publicly and freely available on-line through the MultiDark database
  (www.multidark.org). Our mock catalogs are based on density profile shapes
  that represent the average cluster population of cool-core and non cool-core
  systems (while we account for scatter in the gas mass fractions, i.e., the
  normalization of the profiles). As a result, our model does not fully exploit
  the true variance across the average profile shapes which may result in a
  limited variance of thermal cluster observables.  Hence, the mock catalogs
  will allow quantitative comparisons to future hydrodynamical simulations and
  forecasts for future surveys.  We emphasize that the catalogs are not suited
  for detailed cosmological studies in their (present) published form, which do
  not address any survey response functions nor provide light-cone mock
  catalogs. However, future surveys should be able to test the underlying
  assumptions of our modeling approach, which is based on X-ray observations of
  clusters at low redshift ($z<0.2$) and adopts self-similar redshift evolution
  for extrapolating to higher redshifts. We hope that the community can make
  valuable use of these catalogs in synergy with the future radio, X-ray and
  gamma-ray data.}
